\textbf{ Forecast time-scale determines which processes to model }

### 2019-04-25 doc/manuscript.tex
even more ms additions

In the previous example, evolutionary change was the slow process that determined how quickly the combined model shifted from a temporal to spatial forecast. The process of compositional turnover--change in the abundances and identities of species interacting at a site--can have a similar effect at the community level. We demonstrate this using a metacommunity model developed by Alexander et al. (ref). The model describes Lotka-Volterra competitive interactions among plants within sites that are arrayed along an elevation and temperature gradient. Composition varies along the gradient because of a trade-off between growth rate and cold tolerance: cold sites are dominated by slow-growing species that can tolerate low temperatures, while warm sites are dominated by fast-growing species that are cold intolerant. Multiple species can coexist within sites because all species experience stronger competition from conspecifics than from heterospecifics. Sites are linked by dispersal: a specified fraction of offspring of each species leaves the site where they were produced and reaches all other sites with equal probability.

As in our eco-evolutionary example, we simulate a baseline period with variable but stationary temperature, followed by a period of rapid temperature increase and then a final period of stationary temperature. Interannual variation in temperature is the same at all sites, but mean temperature varies among sites. All sites experienced the same increase in mean temperature. We focus on two response variables: 1) total community biomass and 2) the biomass of one focal species that dominates the baseline period at the site in the middle of the gradient.

Dynamics during the baseline period reveal very strong spatial patterns across the mean temperature gradient contrasted by weak responses to interannual variation in temperature at one site. Individual species show classic ``Whittaker" patterns of abundances across the gradient (Fig. \ref{fig:community-composition}). Time-averaged total community biomass increases strongly with temperature, because the fast-growing species at the warm sites produce more biomass, while the average biomass of our focal species shows a classic unimodal pattern across the temperature gradient (Fig. \ref{fig:community-baseline}). These spatial patterns are the basis for our "spatial models" of the temperature-biomass relationship. However, at our focal site in the center of the gradient, both total biomass and the biomass of the focal species are quite insensitive to interannnual variation in temperature (Fig. \ref{fig:community-baseline}). Our "temporal models" estimate this weak temperature effect, along with strong effects of biomass in the previous year (e.g., lags).

@@

We can forecast the effect of temperature increase on total biomass and the focal species' biomass using the spatial models and the temporal models. Once again, we also fit a combined model as a weighted average of the spatial and temporal model, with the weights changing over time. For total biomass, the spatial and temporal models both predict immediate increases in biomass with warming, because all species respond positively to temperature increases. However, over time, as faster growing species increase in abundance at the focal site, biomass continues to increase, approaching the predictions of the spatial model (Fig. \ref{fig:community-forecasts} middle). The predictions of the spatial and temporal model for focal species biomass contrast more dramatically, with the temporal model predicting a large increase in biomass and the spatial model predicting a decrease. Initially, the simulated abundances follow the increase predicted by the temporal model, but as faster-growing species increase in abundance, the biomass of the focal species then decreases, eventually falling below its baseline level (Fig. \ref{fig:community-forecasts} bottom). For total biomass (the sum across all species), the weights for the combined model shift rapidly to favor the spatial model, whereas for focal species biomass, the weights shift more slowly towards the spatial model, as the biomass of the focal species decreases and its competitors' biomass increase (Fig. \ref{fig:community-weights}).

>>>>

In the previous example, evolutionary change was the slow process that determined how quickly the combined model shifted from a temporal to spatial forecast. The process of compositional turnover--change in the abundances and identities of species interacting at a site--can have a similar effect at the level of communities and even ecosystems. 
We demonstrate this using a metacommunity model developed by Alexander et al. (ref). The model describes Lotka-Volterra competitive interactions among plants within sites that are arrayed along an elevation and temperature gradient. Composition varies along the gradient because of a trade-off between growth rate and cold tolerance: cold sites are dominated by slow-growing species that can tolerate low temperatures, while warm sites are dominated by fast-growing species that are cold intolerant. Multiple species can coexist within sites because all species experience stronger competition from conspecifics than from heterospecifics. Sites are linked by dispersal: a specified fraction of offspring of each species leaves the site where they were produced and reaches all other sites with equal probability.

As in our eco-evolutionary example, we simulate a baseline period with variable but stationary temperature, followed by a period of rapid temperature increase and then a final period of stationary temperature. Interannual variation in temperature is the same at all sites, but mean temperature varies among sites. All sites experienced the same increase in mean temperature. We focus on two response variables at one site in the middle of the gradient. : 1) total community biomass and 2) the biomass of one focal species that dominates the baseline period

Dynamics during the baseline period reveal very strong spatial patterns across the mean temperature gradient contrasted by weak responses to interannual variation in temperature at one site. Individual species show classic ``Whittaker" patterns of abundances across the gradient (Fig. \ref{fig:community-composition}). Time-averaged total community biomass increases strongly with temperature, because the fast-growing species at the warm sites produce more biomass, while the average biomass of our focal species shows a classic unimodal pattern across the temperature gradient (Fig. \ref{fig:community-baseline}). These spatial patterns are the basis for our ``spatial models" of the temperature-biomass relationship. However, at our focal site in the center of the gradient, both total biomass and the biomass of the focal species are quite insensitive to interannnual variation in temperature (Fig. \ref{fig:community-baseline}). Our ``temporal models" estimate this weak temperature effect, along with strong effects of biomass in the previous year (e.g., lags).

@@

We can forecast the effect of temperature increase on total biomass and the focal species' biomass using the spatial models and the temporal models. For total biomass, the spatial and temporal models both predict immediate increases in biomass with warming, because all species respond positively to temperature increases. However, over time, as faster growing species increase in abundance at the focal site, biomass continues to increase, approaching the predictions of the spatial model (Fig. \ref{fig:community-forecasts} middle). The predictions of the spatial and temporal model for focal species biomass contrast more dramatically, with the temporal model predicting a large increase in biomass and the spatial model predicting a decrease. Initially, the simulated abundances follow the increase predicted by the temporal model, but as faster-growing species increase in abundance, the biomass of the focal species then decreases, eventually falling below its baseline level (Fig. \ref{fig:community-forecasts} bottom). Once again, we also fit a combined model as a weighted average of the spatial and temporal model, with the weights changing over time. For total biomass (the sum across all species), the weights for the combined model shift rapidly to favor the spatial model, whereas for focal species biomass, the weights shift more slowly towards the spatial model, as the biomass of the focal species decreases and its competitors' biomass increase (Fig. \ref{fig:community-weights}).
